Forcing for Isabelle

### 2018-12-07 LICS/renaming.tex
Explicación conceptual de renaming

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Renaming}

In the course of our work we need to reason about renaming of formulas
and its effect on their satisfiability. Since internalized formulas
are implemented using de Bruijn indices, thus the arity of a formula
$\phi$ gives the least set containing all the free variables in
$\phi$. Following \citet{fiore-abssyn}, one can understand the arity
of a formula as the context of the free variables and renamings are,
consequently, mappings changing the context.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Renaming}
\newcommand{\renaming}[2]{#1[#2]}
\newcommand{\inFm}[2]{\mathsf{Member}(#1,#2)}
\newcommand{\eqFm}[2]{\mathsf{Equal}(#1,#2)}
\newcommand{\nandFm}[2]{\mathsf{Nand}(#1,#2)}
\newcommand{\forallFm}[1]{\mathsf{Forall(#1)}}

In the course of our work we need to reason about renaming of formulas
and its effect on their satisfiability. Internalized formulas are
implemented using de Bruijn indices for variables and the arity of a
formula $\phi$ gives the least natural number containing all the free
variables in $\phi$. Following \citet{fiore-abssyn}, one can
understand the arity of a formula as the context of the free
variables; notice that the arity of $\forallFm{\phi}$ is the
predecessor of the arity of $\phi$. Renamings are, consequently,
mappings between finite sets; since we can think of $\mathsf{succ}(n)$
as the co-product $1+n$, then given a renaming $f \colon n \to m$, the
unique morphism $\mathsf{id}_1+f \colon 1+n \to 1+m$ is used to
rename free variables in a quantified formula:

\begin{definition}[Renaming]
  Let $\phi$ be a formula of arity $n$ and let $f \colon n \to m$, the
  renaming of $\phi$ by $f$, denoted $\renaming{\phi}{f}$, is defined
  by recursion on $\phi$:
  \begin{gather*}
    \renaming{\inFm{i}{j}}{f} = \inFm{f\,i}{f\,j}\\
    \renaming{\eqFm{i}{j}}{f} = \eqFm{f\,i}{f\,j}\\
    \renaming{\nandFm{\phi}{\psi}}{f} = \nandFm{\renaming{\phi}{f}}{\renaming{\psi}{f}}\\
    \renaming{\forallFm{\phi}}{f} = \forallFm{\renaming{\phi}{\mathsf{id}+f}}
  \end{gather*}
\end{definition}

Environments can be thought, then, as mappings from a finite set to
the model $M$; the action of renaming on environments is simply given
by composition.

\begin{lemma}
  Let $\phi$ be a formula of arity $n$, $f \colon n \to m$ be a
  renaming, and let $\rho \colon n \to M$ and $\rho' \colon m \to M$
  be environments. If for all $i \in n$, $\rho\,i = \rho'(f\,i)$, then
  $M,\rho\models \phi$ is equivalent to
  $M,\rho' \models \renaming{\phi}{f}$.
\end{lemma}
